---
---
title: "Exploratory Data Analysis of Mobile Payment Fraud"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# About the data

This dataset contains data of simulated mobile money transactions in an African country.
The dataset was downloaded the datasets from kaggle.com. 
The dataset was found tidy,
so no cleaning was necessary before undergoing the analysis.

# Initial questions

In this project, I will conduct exploratory analysis of the dataset to find
interesting insights and trends that would warrant more in-depth investigation.

Initital questions that I have of the data that will guide me are: 

When/where does fraud occur?
What types of transactions does fraud occur?
Are there any variables that are correlated to fraud?

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(ggplot2)
library(gridExtra)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
setwd("~/Documents/R")
payments <- read.csv('mobilePayments.csv')
```

## Adding variable 'days'

```{r echo=FALSE, days}
payments$days <- (payments$step) / 24
```

The step variable measures each hour in a 30 day period.
Dividing this variable by 24 will make it easier to see the days in which transactions happen.

# Overview

```{r echo=FALSE, overview}
str(payments)
```

The dataset has 6362620 observations and 12 variables. The
variable types are integers, numbers, and factors.

# Investigation

## Number of fraudulent cases

```{r echo=FALSE, fraud_rate}
length(payments$isFraud[payments$isFraud == 1])
mean(payments$isFraud)
mean(payments$isFlaggedFraud)
```

Fraudulent transactions are marked with a "1" in the dataset. There are 8213 
of these transactions in the dataset, giving it a 'fraud rate' of 0.13%.
There is another variable, isFlaggedFraud, which represents
whether a transaction was seen by the company as possibly fraudulent
as it was being made. The flagged transactions are marked
with a "1" as well. All flagged transactions were fraudulent. 
Only sixteen transactions are flagged. 
This means that most of the fraudulent transactions get through the system. 

## Averages

```{r echo=FALSE, mean_values}
# Mean of all transactions
mean(payments$amount)
# Splitting fradulent, flagged, and non-fraudulent transactions
fraud <- payments[payments$isFraud == 1, ]
noFraud <- payments[payments$isFraud == 0, ]
isFlaggedFraud <- payments[payments$isFlaggedFraud == 1, ]
# Mean of fraudulent and non-fraudulent transactions
mean(fraud$amount)
mean(noFraud$amount)
mean(isFlaggedFraud$amount)
# Size difference between fraudulent and non-fraudulent transactions
mean(fraud$amount) / mean(noFraud$amount)
# Size difference between non-flagged and flagged fraudulent transactions
mean(isFlaggedFraud$amount) / mean(fraud$amount)
```

Transactions that are fraudulent are more than
eight times larger than legitimate ones, 
but since there are so many more legitimate transactions,
the difference in size barely changes the overall average.

The fraudulent transactions that were flagged are three times larger than the 
ones that weren't. 
This could be because the system flags large transactions in 
excess of 200,000 of the local currency.
Smaller transactions that are still fraudulent are not flagged.

## Average boxplot and histograms

```{r echo=FALSE, message=FALSE, warning=FALSE, amountPlots}

# Boxplot of transactions
ggplot(payments, aes('', amount)) + 
  geom_boxplot()

# Histogram of transactions
ggplot(payments, aes(amount)) + 
  geom_histogram(bins = 100) + 
  xlim(0, 2000000)

```

The data are heavily skewed to the right, with many outliers representing
large transactions.

```{r echo=FALSE, message=FALSE, warning=FALSE, fraudPlots}

# Boxplot
ggplot(fraud, aes('', amount)) + geom_boxplot()

# Histogram
ggplot(fraud, aes(amount)) + geom_histogram(bins = 100)


```

Compared with the whole dataset, the distribution
of the fraudulent transactions is similar, but
there is a spike in the number of transactions around the one million amount.

## Top fraudsters

```{r echo=FALSE, message=FALSE, warning=FALSE, topFraudsters}
# Adding the fraudulent transaction amounts by destination account
group <- aggregate(fraud$amount, by = list(fraud$nameDest), FUN = sum)
# Ordering the group dataframe by transaction amount
group.ordered <- group[order(group$x, decreasing = TRUE), ]
# Top 10
head(group.ordered, 10)
```

One intrepid individual made fraudulent transactions of
over 10 million of the local currency.

## Looking at an invididual fraudster

```{r echo=FALSE, message=FALSE, warning=FALSE, individual}

payments[payments$nameDest == 'C668046170', ]

```

There are five transactions and
three different types of transactions done by this person.
The transactions are distributed throughout the month.
Two of the transactions are fraudulent, and none were flagged as fraudulent.

## Which transaction types have the most fraud

### Transfers

```{r echo=FALSE, message=FALSE, warning=FALSE, transfers}

# Transfers
nrow(fraud[fraud$type == 'TRANSFER', ])
```

### Payments

```{r echo=FALSE, message=FALSE, warning=FALSE, payments}
nrow(fraud[fraud$type == 'PAYMENT', ])
```

### Debits

```{r echo=FALSE, message=FALSE, warning=FALSE, debits}
nrow(fraud[fraud$type == 'DEBIT', ])
```

### Cash in

```{r echo=FALSE, message=FALSE, warning=FALSE, cashIn}
nrow(fraud[fraud$type == 'CASH_IN', ])
```

### Cash out

```{r echo=FALSE, message=FALSE, warning=FALSE, cashOut}
nrow(fraud[fraud$type == 'CASH_OUT', ])
```

Transfers and cashouts have an almost equal number of fraudulent transactions. 
This makes sense, because the dataset documentation says that a common fraud
scheme used is to steal money by transferring it to another account
and then cashing out.

## Dates of transactions

```{r echo=FALSE, message=FALSE, warning=FALSE, dates}

# Histogram of dates

ggplot(payments, aes(days)) +
  geom_histogram()
```


Around day two there is the mode of the graph,
and around day 17 there is a sharp decline in the number of transactions.
A hypothesis about why this pattern occurs is because
people receive their wages at the beginning of the month
and are thus more able to pay for things.
At the end of the month they don't buy as much
because they don't have as much money.

The fraudulent transactions are more normally distributed. 

Put together, this means that a greater percentage of transactions 
later in the month are fraudulent than in the beginning of the month.

# Discussion

Interesting findings so far are:
the difference in distributions of days
on which the fraudulent transactions happen,
and that the fraudulent transactions in this dataset
are larger than legitimate ones.
It is a good sign that so
few transactions are fraudulent, 
but also worrying that so few fraudulent
transactions were flagged, which could suggest inadequate detection practices.

To improve detection of fraud,
it could be useful to place greater scrutiny
on transactions on large transactions,
transactions that happen later in the month,
as well as to more closely monitor or block accounts that have been flagged,
so that they do not make multiple transactions.